---
title: "project1"
author: "Shanyao Jing"
date: "2/21/2020"
output: html_document
---
library(stringr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyr)
library(tibble)
library(ggthemes)
library(maps)

#Clean Data
setwd("C:/Users/ShineJean/Documents/2019/0-2020Spring/IE6600/Project1")
data <- read.csv("fmarket.csv", 
                 encoding="UTF-8", stringsAsFactors = F, na.strings = c(""))
data$ts<-as.character(data$updateTime)
data$year<-str_extract(data$ts, "20+[:digit:]+[:digit:]")
#write.csv(data, 'fmarket_clean.csv')

#Examine
#ts<-select(data,updateTime,year)
#write.csv(ts, 'ts.csv')

count<-data%>%
  group_by(State,year) %>%
  summarise(Count=n())

```{r}
#map
#Figure1
y<-data%>%
  group_by(State) %>%
  summarise(Count=n())
us_states <- map_data('state')
y$State <- tolower(y$State)
mode(y$Count)
colnames(y)[1]='region'
us_states_fmarket <- left_join(us_states,y)
head(us_states_fmarket)
mode(us_states_fmarket$Count)

snames <- aggregate(cbind(long,lat)~region,data=us_states_fmarket,FUN=function(x) mean(range(x)))
ggplot()+
  geom_polygon(data = us_states_fmarket,
               mapping = aes(x = long, y = lat, group = group, fill = Count),
               color = "gray90", size = 0.1)+
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradient(name='Count',low='white',high='#840050')+
  labs(title = "Number of Markets across USA") + 
  theme_map()+
  labs(fill = "count")+ 
  geom_text(data=snames,aes(long,lat,label=region))+
  theme_economist(base_size = 14)+
  theme(legend.text=element_text(size=rel(0.6)))+
  xlab("Longitude")+ylab("Latitude")

ggsave("map.png",device = "png",width = 12, height = 12, dpi = 300)
 
```

  
sc<-data%>%
  group_by(State) %>%
  summarise(Count=n()) 
count2<-filter(count, State == sc[1,1])
count2<- as.data.frame(count2[,-1])

for (i in 2:53) {
  df <- filter(count, State == sc[i,1])
  df <- as.data.frame(df[,-1])
  count2 <- merge(count2,df,by="year", all = TRUE)
}
 
 
x<-pull(sc,State)
colnames(count2)[2:54]<-x
count_t<-data.frame(t(count2),stringsAsFactors = F)
sum(is.na(count_t))
count_t[is.na(count_t)]<-0

start<-data.frame(count_t[-1,1])
colnames(start)="origin"

#growth rate
growth<-data%>%
  group_by(State) %>%
  summarise(Count=n())
growth<-cbind(growth,start)


a<-data.frame(growth[,2],stringsAsFactors = F)
b<-data.frame(growth[,3],stringsAsFactors = F)
b<-as.data.frame(lapply(b,as.numeric))
c<-(a-b)/11

growth<-growth%>%
cbind(as.data.frame(round(c,2),stringsAsFactors = F))
colnames(growth)[4]<-"growth_rate"

```{r}
#Figure2
rate_T10_1 <- growth %>%
  arrange(desc(origin)) %>%
  head(10)
b<-data.frame(rate_T10_1[,3],stringsAsFactors = F)
b<-as.data.frame(lapply(b,as.numeric))
rate_T10_1<-rate_T10_1%>%
cbind(b,stringsAsFactors = F)
colnames(rate_T10_1)[5]<-"origin2"

ggplot(data=rate_T10_1, aes(x = reorder(State, origin2), y = origin2, fill = origin2))+
  geom_bar(stat = "identity")+coord_flip()+
  xlab("Top 10 States")+ylab("Numbers of Markets in 2009")+
  labs(fill = "Number of Markets")+
  geom_text(data=rate_T10_1,aes(x=State,y=10,label=origin2),color="white")+
  theme_economist(base_size = 14)+
  ggtitle("Top 10 States of Markests in 2009")+
  scale_fill_gradient(low='#7ad2f6',high='#014d64')+
  ylim(0,200)+
  theme(legend.text=element_text(size=rel(0.8)))
  
ggsave("2009.png",device = "png",width = 10, height = 7, dpi = 300)

```



```{r}
#Figure4
rate_T10_2 <- growth %>%
  arrange(desc(growth_rate)) %>%
  head(10)

ggplot(data=rate_T10_2, aes(x = reorder(State, growth_rate), y = growth_rate, fill = growth_rate))+
  geom_bar(stat = "identity")+coord_flip()+
  xlab("Top 10 States")+ylab("Average Annual Increasing Numbers of Markets")+
  labs(fill = "number of markets")+
  #scale_fill_gradient(low='blue',high='red')+
  geom_text(data=rate_T10_2,aes(x=State,y=5,label=growth_rate),color="white")+
  theme_economist(base_size = 14)+
  ggtitle("Top 10 States of Markets Growth from 2009 to 2020")+
  scale_fill_gradient(low='#7ad2f6',high='#014d64')+
  theme(legend.text=element_text(size=rel(0.8)))+
  

ggsave("2009-2020.png",device = "png",width = 10, height = 7, dpi = 300)

```


```{r}
#Figure3
rate_T10_3 <- growth %>%
  arrange(desc(Count)) %>%
  head(10)

ggplot(data=rate_T10_3, aes(x = reorder(State, Count), y = Count, fill = Count))+
  geom_bar(stat = "identity")+coord_flip()+
  xlab("Top 10 States")+ylab("Numbers of markets in 2020")+
  labs(fill = "number of markets")+
  geom_text(data=rate_T10_3,aes(x=State,y=50,label=Count),color="white")+
  theme_economist(base_size = 14)+
  ggtitle("Top 10 States of Markests in 2020")+
  scale_fill_gradient(low='#7ad2f6',high='#014d64')+
  theme(legend.text=element_text(size=rel(0.6)))+
  ylim(0,800)

ggsave("2020.png",device = "png",width = 10, height = 7, dpi = 300)

```

```{r}
#view column names
colnames(data)
data%>%
```

```{r}
#Product
#Figure5
product<-data%>%
  select(State,Organic,Bakedgoods,Cheese,Crafts,Flowers,Eggs,Seafood,Herbs,Vegetables,Honey,Jams,Maple,Meat,Nursery,Nuts,Plants,Poultry,Prepared,Soap,Trees,Wine,Coffee,Beans,Fruits,Grains,Juices,Mushrooms,PetFood,Tofu,WildHarvested)

product[product == 'Y'] = 1
product[product == 'N'] = 0
product[product == '-'] = 0

product2<-as.data.frame(lapply(product[,-1],as.numeric))    
state<-data.frame(product[,1],stringsAsFactors = F)
product<-cbind(state,product2)
colnames(product)[1]<-"State"

NorthEast<-c("Maine","New Hampshire","Vermont","Massachusetts","Rhode Island","Connecticut",
            "New York","New Jersey","Pennsylvania")

MidWest<-c("Ohio", "Michigan", "Indiana", "Wisconsin", "Illinois", "Minnesota", "Iowa",
           "Missouri", "North Dakota", "South Dakota", "Nebraska", "Kansas")

South<-c("Delaware","Maryland","District of Columbia","Virginia","West Virginia",
         "North Carolina","South Carolina","Georgia","Florida","Kentucky","Tennessee",
         "Mississippi","Alabama","Oklahoma","Texas","Arkansas","Louisiana")

West<-c("Idaho","Montana","Wyoming","Nevada","Utah","Colorado","Arizona","New Mexico",
        "Alaska","Washington","Oregon","California","Hawaii")

product$region<-ifelse(product$State %in% NorthEast,"NorthEast",
       ifelse((product$State %in% MidWest),"MidWest",
            ifelse((product$State %in% South),"South",
               ifelse((product$State %in% West),"West","Other"))))

pr<-product %>%
  group_by(region) %>%
  summarise(Count=n()) 
pr<-pr[,-2]

tn<-data.frame(c("Organic","Bakedgoods","Cheese","Crafts","Flowers","Eggs","Seafood","Herbs","Vegetables","Honey","Jams","Maple","Meat","Nursery","Nuts","Plants","Poultry","Prepared","Soap","Trees","Wine","Coffee","Beans","Fruits","Grains","Juices","Mushrooms","PetFood","Tofu","WildHarvested"))

for (t in 1:30) {
  a<-tn[t,1]
  filter<-filter(product,product[,t+1]==1)
  filter<-filter[,c(t+1,32)]
  type<-filter %>%
  group_by(region) %>%
  summarise(Count=n()) 
  pr<-merge(pr,type,by = "region", all = TRUE)
 
}


pr[is.na(pr)] <- 0
colnames(pr)[-1]<-c("Organic","Bakedgoods","Cheese","Crafts","Flowers","Eggs","Seafood","Herbs","Vegetables","Honey","Jams","Maple","Meat","Nursery","Nuts","Plants","Poultry","Prepared","Soap","Trees","Wine","Coffee","Beans","Fruits","Grains","Juices","Mushrooms","PetFood","Tofu","WildHarvested")
prt<-data.frame(t(pr),stringsAsFactors = F)
head（prt)
prt<-prt[-1,]
colnames(prt) <- c("MidWest","NorthEast","Other","South","West")
prt<- rownames_to_column(prt, var= "region") 
prtg<-as.data.frame(gather(prt,distribution,count,-region))
mode(prtg$count)
prtg$count<-as.numeric(prtg$count)

label1<-summarise(group_by(prtg,region),amount=sum(count))
prtg<-left_join(prtg,label1)
#write.csv(prtg,"bar.csv")

ggplot(data=prtg, aes(x=reorder(region,count),y=count,fill=distribution))+
  geom_bar(stat = "identity",position="stack")+xlab("Types of Products")+coord_flip()+
  ylab("Numbers of Markets")+
  labs(fill = "Regions")+
  theme_economist(base_size = 14)+scale_fill_economist()+
  ggtitle("Distribution of Products by Regions")+
  theme(legend.text=element_text(size=rel(0.6)))+
  geom_text(data=prtg,aes(x=reorder(region,count),y=6000,label=amount)，color="#808080")
  #theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
  



ggsave("product.png",device = "png",width = 10, height = 10, dpi = 300)
```

```{r}
#Pie plot
#Figure6
pie<-data%>%
  select(State, Credit, WIC, WICcash, SFMNP, SNAP)

pie[pie == 'Y'] = 1
pie[pie == 'N'] = 0
pie[pie == '-'] = 0

pie2<-as.data.frame(lapply(pie[,-1],as.numeric))    
state2<-data.frame(pie[,1],stringsAsFactors = F)
pie<-cbind(state2,pie2)
colnames(pie)[1]<-"State"

NorthEast<-c("Maine","New Hampshire","Vermont","Massachusetts","Rhode Island","Connecticut",
            "New York","New Jersey","Pennsylvania")

MidWest<-c("Ohio", "Michigan", "Indiana", "Wisconsin", "Illinois", "Minnesota", "Iowa",
           "Missouri", "North Dakota", "South Dakota", "Nebraska", "Kansas")

South<-c("Delaware","Maryland","District of Columbia","Virginia","West Virginia",
         "North Carolina","South Carolina","Georgia","Florida","Kentucky","Tennessee",
         "Mississippi","Alabama","Oklahoma","Texas","Arkansas","Louisiana")

West<-c("Idaho","Montana","Wyoming","Nevada","Utah","Colorado","Arizona","New Mexico",
        "Alaska","Washington","Oregon","California","Hawaii")

pie$region<-ifelse(pie$State %in% NorthEast,"NorthEast",
       ifelse((pie$State %in% MidWest),"MidWest",
            ifelse((pie$State %in% South),"South",
               ifelse((pie$State %in% West),"West","Other"))))

           
P_count<-data.frame(colSums(pie[,c(2:6)]),stringsAsFactors = F)
colnames(P_count)<-"count"
amount<-colSums(P_count['count'])
P_count<-rownames_to_column(P_count,var="Payment")


P_count<-mutate(P_count,percentage=round(count/amount*100,1))

#write.csv(P_count, 'P_count.csv')
  
label<-paste(P_count$Payment,P_count$count,P_count$percentage)
label<-paste(label,"%",sep="")
number<-P_count$percentage
pie(number,labels =label,col=rainbow(length(label)),main="Percentage of Payment Method across USA")


ggplot(data=P_count, aes(x = "", y = percentage, fill = percentage))+
geom_bar(stat = "identity")+
coord_polar("y", start=0)+
theme_economist(base_size = 14)+
geom_text(aes(y = P_count$percentage/2 +
  c(0,cumsum(P_count$percentage)[-length(P_count$percentage)]), x =
  sum(P_count$percentage)/90, label = label),color="white")+
  ggtitle("Percentage of Payment System")+
theme(legend.text=element_text(size=rel(0.6)))
 

ggsave("payment.png",device = "png",width = 8, height = 8, dpi = 300)
 
```


```{r}
#Figure11
fmarket<-data
f1 <-
  fmarket %>% 
  filter(Fruits=='Y') %>% 
  group_by(State) %>% 
  summarise(count=n())
us_states <- map_data('state')
f1$State <- tolower(f1$State)
colnames(f1)[1]='region'
us_states_fmarket <- left_join(us_states,f1)
head(us_states_fmarket)
snames <- aggregate(cbind(long,lat)~region,data=us_states_fmarket,FUN=function(x) mean(range(x)))
ggplot()+
  geom_polygon(data = us_states_fmarket,
               mapping = aes(x = long, y = lat, group = group, fill = count),
               color = "gray90", size = 0.1)+
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradient(name='Count',low='white',high='#50b883')+
  labs(title = "Distribution of Fruits across USA") + 
  theme_map() + 
  labs(fill = "count")+
  theme(legend.text=element_text(size=14))+ 
  geom_text(data=snames,aes(long,lat,label=region))+
  theme_economist(base_size = 14)+
  theme(legend.text=element_text(size=rel(0.6)))+
  xlab("Longitude")+ylab("Latitude")

ggsave("map1.png",device = "png",width = 12, height = 12, dpi = 300)
 
```

```{r}
#Figure10
f2 <-
  fmarket %>% 
  filter(fmarket$SFMNP=='Y') %>% 
  group_by(State) %>% 
  summarise(count=n())
us_states <- map_data('state')
f2$State <- tolower(f2$State)
colnames(f2)[1]='region'
us_states_fmarket <- left_join(us_states,f2)
us_states_fmarket[is.na(us_states_fmarket)] <- 0
ggplot()+
  geom_polygon(data = us_states_fmarket,
               mapping = aes(x = long, y = lat, group = group, fill = count),
               color = "gray90", size = 0.1)+
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradient(name='Count',low='white',high='orange')+
  labs(title = "Distribution of SFMNP in Each State")+
  theme_map() + 
  labs(fill = "count")+ 
  theme(legend.text=element_text(size=8))+ 
  geom_text(data=snames,aes(long,lat,label=region)) +
  theme_economist(base_size = 14)+
  theme(legend.text=element_text(size=rel(0.6)))+
  xlab("Longitude")+ylab("Latitude")

ggsave("map2.png",device = "png",width = 12, height = 12, dpi = 300)
 
```


```{r}
#line
#Figure8
fmarket<-data
d_fruits <- fmarket %>% filter(Fruits=='Y') %>% 
  filter(SFMNP=='Y') %>% select(year)
#d_f <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=Freq+lag(Freq)) %>% select(-Freq)
d_1 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_1
colnames(d_1)=c('year','Fruit')
d_fruits <- fmarket %>% filter(Vegetables=='Y') %>% 
  filter(SFMNP=='Y') %>% select(year)
d_2 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_2
colnames(d_2)=c('year','Vegetables')
d_fruits <- fmarket %>% filter(Honey=='Y') %>% 
  filter(SFMNP=='Y') %>% select(year)
d_3 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_3
colnames(d_3)=c('year','Honey')
d_fruits <- fmarket %>% filter(Herbs=='Y') %>% 
  filter(SFMNP=='Y') %>% select(year)
d_4 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_4
colnames(d_4)=c('year','Herbs')
d_all <- left_join(left_join(left_join(d_2,d_3),d_4),d_1)
d_all$year <- as.numeric(d_all$year)
d_all[1,5]=0
d_all[2,5]=0
df <- d_all %>% 
  tidyr::gather(key='variable',value='value',-year)
df
df[is.na(df)] <- 0
ggplot(df, aes(x = year, y = value)) + 
  geom_line(size=1,aes(color = variable))+
  geom_point(size=3, shape=20,color='orange')+
  scale_x_discrete(limits=2011:2020)+scale_color_manual(values = c("#840050",'#8bc3a0','#36345c','#895617'))+
  theme_bw()+
  labs(xlab = "Year", ylab = "Count",
  title = "Market Growth of Products Supported by SFMNP")+
  theme_economist(base_size = 14)+
  theme(legend.title = element_blank(),legend.text=element_text(size=rel(1)))+
  xlab("Year")+ylab("Number of Markets")+
  ylim(0,3000)

ggsave("line.png",device = "png",width =9, height = 10, dpi = 300)
 
```

```{r}
#line2
#Figure9
fmarket<-data
d_fruits <- fmarket %>% filter(Fruits=='Y') %>% 
  filter(SFMNP!='Y') %>% select(year)
d_1 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_1
colnames(d_1)=c('year','Fruit')
d_fruits <- fmarket %>% filter(Vegetables=='Y') %>% 
  filter(SFMNP!='Y') %>% select(year)
d_2 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_2
colnames(d_2)=c('year','Vegetables')
d_fruits <- fmarket %>% filter(Honey=='Y') %>% 
  filter(SFMNP!='Y') %>% select(year)
d_3 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_3
colnames(d_3)=c('year','Honey')
d_fruits <- fmarket %>% filter(Herbs=='Y') %>% 
  filter(SFMNP!='Y') %>% select(year)
d_4 <- table(d_fruits) %>% as.data.frame() %>% mutate(Real=cumsum(Freq))%>% select(-Freq)
d_4
colnames(d_4)=c('year','Herbs')
d_all <- left_join(left_join(left_join(d_2,d_3),d_4),d_1)
d_all$year <- as.numeric(d_all$year)
d_all[1,5]=0
d_all[2,5]=0
df <- d_all %>% 
  tidyr::gather(key='variable',value='value',-year)
df
df[is.na(df)] <- 0
ggplot(df, aes(x = year, y = value)) + 
  geom_line(size=1,aes(color = variable))+
  geom_point(size=3, shape=20,color='orange')+
  scale_x_discrete(limits=2011:2020)+scale_color_manual(values = c("#840050",'#8bc3a0','#36345c','#895617'))+
  theme_bw()+
  labs(xlab = "Year", ylab = "Count",
  title = "Market Growth of Products Not Supported by SFMNP")+
  theme_economist(base_size = 14)+
  theme(legend.title = element_blank(),legend.text=element_text(size=rel(1)))+
  xlab("Year")+ylab("Number of Markets")

ggsave("line2.png",device = "png",width = 9, height = 10, dpi = 300)
 
```

